---
title: "2023-Assessement"
output: html_document
date: "2024-08-27"
---

Import R librarys and weather data.  

```{r}
library(data.table)
library(viticolaR)
library(epiphytoolR)
library(weatherOz)
library(ggplot2)
library(nasapower)

# read in north Tamborine weather data
if(file.exists("./data/23-24_NTamborine.csv") == FALSE){
   download.file(url = "https://filedn.eu/lKw35gljYV2BIxxlGg9SUJb/weather/23-24_NTamborine.csv",
                 destfile = "./data/23-24_NTamborine.csv")
}
ntw <- fread("./data/23-24_NTamborine.csv")

if(dir.exists("./data/") == FALSE) dir.create("./data")
if(dir.exists("./cache/") == FALSE) dir.create("./cache")
```


Inspect weather data.  

```{r}
head(ntw)[,1:10]
```

Some columns need formatting

```{r}
# Format time record
ntw[,time_utc := as.POSIXct(as.character(aifstime_utc), format= "%Y%m%d%H%M%S")]
ntw[,time_local := as.POSIXct(as.character(aifstime_local), format= "%Y%m%d%H%M%S", tz = "Australia/Brisbane")]

# place in order
ntw <- ntw[order(time_utc)]
```

Check data for duplication

```{r}
any(duplicated(ntw))
```

```{r}
#lonitude was slightly variable
range(ntw$lon)
range(ntw$lat)
ntw[,lon := mean(lon)]
ntw[,lat := mean(lat)]
```


Slice data to season
```{r}
ntw <- ntw[time_utc > as.POSIXct("2023-06-30", tz = "UTC")]
```



Format the weather data for a model readable version

```{r}
ntw_c <- format_weather(ntw,
                        POSIXct_time = "time_utc", 
                        time_zone = "UTC",
                        temp = "air_temp",
                        rain = "rain_ten",
                        rh = "rel_hum",
                        ws = "wind_spd_kmh",
                        wd = "wind_dir_deg",
                        station = "name",
                        lon = "lon",
                        lat = "lat",
                        data_check = FALSE,
                        impute_nas = FALSE
                        )
```

This data has gaps Lets inspect the coverage  
```{r}
ntw_c[is.na(rain)]
ntw_c[is.na(temp)]
ntw_c[is.na(rh)]
```

This shows there is a fair bit of missing data. 
The model does not handle missing data, we can fill missing data from two sources,
remotely sensed weather from nasapower satelites and/or the openmet package.  

### nasapower suplemented weather
Get silo weather data to patch NA data

```{r,cache=TRUE}
if(file.exists("cache/23-24_weatherdata_nasapower.csv")){
   nasa_w <- fread("cache/23-24_weatherdata_nasapower.csv")
}else{
   nasa_w <- nasapower::get_power(community = "AG",
                                  temporal_api = "hourly",
                                  lonlat = c(ntw_c$lon[1],ntw_c$lat[1]),
                                  pars = c("RH2M", "T2M", "PRECTOTCORR"),
                                  dates = c("2023-07-01", "2024-03-30"))
   
   setDT(nasa_w)
   
   # format a posixct column
   nasa_w[,times := as.POSIXct(paste0(YEAR,"-",MO,"-",DY," ",HR,":00:00"))]
   
   # save data
   fwrite(nasa_w,file = "cache/23-24_weatherdata_nasapower.csv")
}
```


Merge weather station data from field site and nasapower weather data.  

```{r}
ntw_n <- merge.data.table(ntw_c,nasa_w,by = "times")
```

Fill missing or odd values data with `nasapower` data.

```{r correct_with_nasa}
# Fill missing data with nasapower data
ntw_n[is.na(temp),temp := T2M]
ntw_n[is.na(rh),rh := RH2M]
ntw_n[is.na(rain),rain := PRECTOTCORR]

# replace values outside range with NASA POWER data
ntw_n[temp <= -30, temp := T2M]
ntw_n[temp >= 60, temp := T2M]
ntw_n[rh < 0, rh := RH2M]

# format again and just check temperature relative humidity and rainfall
# use data_check to ensure all looks good for the data needed for the model
ntw_n <- format_weather(ntw_n,
                        time_zone = "UTC",
                        data_check = c("temp","rh","rain"))

```


#### Run model nasa  

```{r}
mod_dm <- viticolaR::estimate_DM_PI(ntw_n,
                                    Start = as.POSIXct("2023-07-01"),
                                    End = "2024-03-30")
# show summary
summary(mod_dm)
```


```{r}
# Get the Primary infection dates
downy_dates <- get_PI_dates(mod_dm)
```

Field observations indicated that downy mildew was first detected on the 10th of 
November 2023.  
```{r}
DM_dectected <- as.POSIXct("2023-11-10")
```


Plot the hydrothermal time for the 2023 season.

```{r}
data.frame(time =  mod_dm$time_hours[mod_dm$Hyd_t < 15],
           hydrotime = mod_dm$Hyd_t[mod_dm$Hyd_t < 15])|>
   ggplot(aes(x = time, hydrotime)) +
   geom_line()+
   geom_rect(aes(xmin = mod_dm$time_hours[1],
                 xmax = mod_dm$time_hours[250*24],
                 ymin = 1.3,
                 ymax = 8.6),
             fill = "skyblue", alpha =0.01)+
   theme_minimal()+
   theme(axis.text.x = element_text(angle = 90, 
                                            vjust = 0.5, 
                                            hjust=1))+
   ylim(c(0,15))+
   scale_x_continuous(breaks = (seq(min(mod_dm$time_hours)+10*60*60,
                                    max(mod_dm$time_hours) - 60*60*24*7*4,
                                    by = 60*60*24*7)))+
   geom_vline(xintercept = DM_dectected, colour = "red")+
   ylab("Hydrothermal time")

```


```{r,cache=TRUE, fig.dim=c(30,10)}
SPO_plot <-
   ggplot() +
         geom_ribbon_viticolaR(mod_dm)+
         geom_line_viticolaR(mod_dm)+
         scale_fill_gradient(name = "Mature Sporangia\ncohorts",
                             low = "#EBE9CF",
                             high = "#ADA205")+
         scale_color_continuous(name = "Immature sporangia\ncohorts")+
         theme_minimal()+
         coord_cartesian(ylim = c(0,1.2))+
         ylab("Progress towards sporangia maturity")+
         theme(legend.position="bottom")+
         geom_rect(aes(xmin = head(mod_dm$time_hours,n = 1),
                       xmax = tail(mod_dm$time_hours,n = 1),
                       ymin = 1,
                       ymax = 2),
                       fill = "grey",
                       alpha = 0.6)+
         scale_x_continuous(breaks = seq(min(mod_dm$time_hours),
                                                 max(mod_dm$time_hours),
                                                 by = 60*60*24*7))+
         geom_vline(xintercept = DM_dectected, colour = "red")+
         theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

SPO_plot
ggsave(filename = "cache/SPO_plot.png",
       plot = SPO_plot,
       width = 3000,height = 500, units = "px", dpi = 90)
```




```{r}
# downy_dates <- fread("cache/PI_dates.csv")
# downy_dates <- get_PI_dates(mod_dm)
```

Not sure why get_PI_dates is  not working saved data.table and can read back in.  


```{r}
downy_dates[primary_infection_stage == "GEO_h"][cohort < 55]
downy_dates[primary_infection_stage == "SUS_death_ind"][cohort < 55]
downy_dates[primary_infection_stage == "ZRE_ind"][cohort < 55]
downy_dates[primary_infection_stage == "ZDI_ind"][cohort < 55]
downy_dates[primary_infection_stage == "ZIN_ind"][cohort < 55]
downy_dates[primary_infection_stage == "SUZ_death_ind"][cohort < 55]

downy_dates[primary_infection_stage == "INC_h_upper"][cohort < 55]
```

### openweather  
#### Run model openweather  

```{r}
source("R/test_open-meteo.R")
ntw_o <- tamof[is.na(times) == FALSE]
ntw_o[is.na(temp),temp := hourly_temperature_2m]
ntw_o[is.na(rh),rh := hourly_relative_humidity_2m]
ntw_o[is.na(rain),rain := hourly_rain]
```


```{r}
# replace values outside range with NASA POWER data
ntw_o[temp <= -30, temp := hourly_temperature_2m]
ntw_o[rh < 0, rh := hourly_relative_humidity_2m]

# format again and just check temperature relative humidity and rainfall
ntw_o <- format_weather(ntw_o,
                        time_zone = "UTC",
                        data_check = c("temp","rh","rain"))

```


```{r}
mod_dm <- viticolaR::estimate_DM_PI(ntw_o,
                                    Start = as.POSIXct("2023-07-01"),
                                    End = "2024-03-30")
downy_dates <- get_PI_dates(mod_dm)
fwrite(downy_dates, "cache/PI_dates.csv")
```


```{r}
summary(mod_dm)
```

Field observations indicated that downy mildew was first detected on the 10th of 
November 2023.  
```{r}
DM_dectected <- as.POSIXct("2023-11-10")
```



Plot the hydrothemal time for the 2023 season.

```{r}
data.frame(time =  mod_dm$time_hours[mod_dm$Hyd_t < 15],
           hydrotime = mod_dm$Hyd_t[mod_dm$Hyd_t < 15])|>
   ggplot(aes(x = time, hydrotime)) +
   geom_line()+
   geom_rect(aes(xmin = mod_dm$time_hours[1],
                 xmax = mod_dm$time_hours[250*24],
                 ymin = 1.3,
                 ymax = 8.6),
             fill = "skyblue", alpha =0.01)+
   theme_minimal()+
   theme(axis.text.x = element_text(angle = 90, 
                                            vjust = 0.5, 
                                            hjust=1))+
   ylim(c(0,15))+
   scale_x_continuous(breaks = (seq(min(mod_dm$time_hours)+10*60*60,
                                    max(mod_dm$time_hours) - 60*60*24*7*4,
                                    by = 60*60*24*7)))+
   geom_vline(xintercept = DM_dectected, colour = "red")+
   ylab("Hydrothermal time")

```




```{r,cache=TRUE, fig.dim=c(30,10)}
SPO_plot <-
   ggplot() +
         geom_ribbon_viticolaR(mod_dm)+
         geom_line_viticolaR(mod_dm)+
         scale_fill_gradient(name = "Mature Sporangia\ncohorts",
                             low = "#EBE9CF",
                             high = "#ADA205")+
         scale_color_continuous(name = "Immature sporangia\ncohorts")+
         theme_minimal()+
         coord_cartesian(ylim = c(0,1.2))+
         ylab("Progress towards sporangia maturity")+
         theme(legend.position="bottom")+
         geom_rect(aes(xmin = head(mod_dm$time_hours,n = 1),
                       xmax = tail(mod_dm$time_hours,n = 1),
                       ymin = 1,
                       ymax = 2),
                       fill = "grey",
                       alpha = 0.8)+
         scale_x_continuous(breaks = seq(min(mod_dm$time_hours),
                                                 max(mod_dm$time_hours),
                                                 by = 60*60*24*7))+
         geom_vline(xintercept = DM_dectected, colour = "red")+
         theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

SPO_plot
ggsave(filename = "cache/SPO_plot.png",
       plot = SPO_plot,
       width = 3000,height = 500, units = "px", dpi = 90)
```




```{r}
#downy_dates <- fread("cache/PI_dates.csv")
# downy_dates <- get_PI_dates(mod_dm)
```

Not sure why get_PI_dates is  not working saved data.table and can read back in.  


```{r}
downy_dates[primary_infection_stage == "GEO_h"][cohort < 55]
downy_dates[primary_infection_stage == "SUS_death_ind"][cohort < 55]
downy_dates[primary_infection_stage == "ZRE_ind"][cohort < 55]
downy_dates[primary_infection_stage == "ZDI_ind"][cohort < 55]
downy_dates[primary_infection_stage == "ZIN_ind"][cohort < 55]
downy_dates[primary_infection_stage == "SUZ_death_ind"][cohort < 55]

downy_dates[primary_infection_stage == "INC_h_upper"][cohort < 55]
```



## 2024 Season
### Experimental design  



### Spray applications

 - 2024-09-04  Miravis  
 - 2024-09-18  Captan  
 - 2024-10-02  EcoCarb Sulphur mix  

