---
title: "2023-Assessement"
output: html_document
date: "2024-08-27"
---

Import R librarys and weather data.  

```{r}
library(data.table)
library(viticolaR)
library(epiphytoolR)
library(weatherOz,lib.loc = 'C:/Users/mel096/AppData/Local/R/win-library/4.4')
library(ggplot2)
library(nasapower)

# read in north tamborine weather data
ntw <- fread("./data/23-24_NTamborine.csv")
```


Inspect weather data.  

```{r}
head(ntw)[,1:10]
```

Some columns need formatting

```{r}
# Format time record
ntw[,time_utc := as.POSIXct(as.character(aifstime_utc), format= "%Y%m%d%H%M%S")]
ntw[,time_local := as.POSIXct(as.character(aifstime_local), format= "%Y%m%d%H%M%S", tz = "Australia/Brisbane")]

# place in order
ntw <- ntw[order(time_utc)]
```

Check data for duplication

```{r}
any(duplicated(ntw))
```

```{r}
#lonitude was slightly variable
range(ntw$lon)
range(ntw$lat)
ntw[,lon := mean(lon)]
ntw[,lat := mean(lat)]
```


Slice data to season
```{r}
ntw <- ntw[time_utc > as.POSIXct("2023-06-30", tz = "UTC")]
```



Format the weather data for a model readable version

```{r}
ntw_c <- format_weather(ntw,
                        POSIXct_time = "time_utc", 
                        time_zone = "UTC",
                        temp = "air_temp",
                        rain = "rain_ten",
                        rh = "rel_hum",
                        ws = "wind_spd_kmh",
                        wd = "wind_dir_deg",
                        station = "name",
                        lon = "lon",
                        lat = "lat",
                        data_check = FALSE,
                        muffle_warnings = TRUE
                        )
```

This data has gaps Lets inspect the coverage

```{r}
ntw_c[is.na(rain)]
```


Get silo weather data to patch NA data

```{r,cache=TRUE}
nasa_w <- nasapower::get_power(community = "AG",
                               temporal_api = "hourly",
                               lonlat = c(ntw_c$lon[1],ntw_c$lat[1]),
                               pars = c("RH2M", "T2M", "PRECTOTCORR"),
                               dates = c("2023-07-01", "2024-03-30"))

setDT(nasa_w)

nasa_w[,times := as.POSIXct(paste0(YEAR,"-",MO,"-",DY," ",HR,":00:00"))]
```


merge weather data

```{r}
ntw_c <- merge.data.table(ntw_c,nasa_w,by = "times")

ntw_c[is.na(temp),temp := T2M]
ntw_c[is.na(rh),rh := RH2M]
ntw_c[is.na(rain),rain := PRECTOTCORR]


fwrite(ntw_c,file = "cache/23-24_weatherdata.csv")
```


```{r}
if(missing(ntw_c)) ntw_c <- fread("cache/23-24_weatherdata.csv")
```

### Run model  

```{r}
mod_dm <- viticolaR::estimate_DM_PI(ntw_c,Start = as.POSIXct("2023-07-01"))
downy_dates <- get_PI_dates(mod_dm)
```


```{r}
str(mod_dm[1:6])
```

Field observations indicated that downy mildew was first detected on the 10th of 
November 2023.  
```{r}
DM_dectected <- as.POSIXct("2023-11-10")
```



Plot the hydorlogical time for the 2023 season.

```{r}
data.frame(time =  mod_dm$time_hours,
           hydrotime = mod_dm$Hyd_t)|>
   ggplot(aes(x = time, hydrotime)) +
   geom_line()+
   geom_rect(aes(xmin = mod_dm$time_hours[1],
                 xmax = mod_dm$time_hours[250*24],
                 ymin = 1.3,
                 ymax = 8.6),
             fill = "skyblue", alpha =0.01)+
   theme_minimal()+
   theme(axis.text.x = element_text(angle = 90, 
                                            vjust = 0.5, 
                                            hjust=1))+
   ylim(c(0,15))+
   scale_x_continuous(breaks = (seq(min(mod_dm$time_hours)+10*60*60,
                                    max(mod_dm$time_hours) - 60*60*24*7*4,
                                    by = 60*60*24*7)))+
   geom_vline(xintercept = DM_dectected, colour = "red")+
   ylab("Hydrothermal time")

```




```{r,cache=TRUE, fig.dim=c(30,10)}
SPO_plot <-
   ggplot() +
         geom_ribbon_viticolaR(mod_dm)+
         geom_line_viticolaR(mod_dm)+
         scale_fill_gradient(name = "Mature Sporangia\ncohorts",
                             low = "#EBE9CF",
                             high = "#ADA205")+
         scale_color_continuous(name = "Immature sporangia\ncohorts")+
         theme_minimal()+
         coord_cartesian(ylim = c(0,1.2))+
         ylab("Progress towards sporangia maturity")+
         theme(legend.position="bottom")+
         geom_rect(aes(xmin = head(mod_dm$time_hours,n = 1),
                       xmax = tail(mod_dm$time_hours,n = 1),
                       ymin = 1,
                       ymax = 2),
                       fill = "grey",
                       alpha = 0.6)+
         scale_x_continuous(breaks = seq(min(mod_dm$time_hours),
                                                 max(mod_dm$time_hours),
                                                 by = 60*60*24*7))+
         geom_vline(xintercept = DM_dectected, colour = "red")+
         theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

SPO_plot
ggsave(filename = "cache/SPO_plot.png",
       plot = SPO_plot,
       width = 3000,height = 500, units = "px", dpi = 90)
```




```{r}
downy_dates <- fread("cache/PI_dates.csv")
# downy_dates <- get_PI_dates(mod_dm)
```

Not sure why get_PI_dates is  not working saved data.table and can read back in.  


```{r}
downy_dates[primary_infection_stage == "GEO_h"][cohort < 55]
downy_dates[primary_infection_stage == "SUS_death_ind"][cohort < 55]
downy_dates[primary_infection_stage == "ZRE_ind"][cohort < 55]
downy_dates[primary_infection_stage == "ZDI_ind"][cohort < 55]
downy_dates[primary_infection_stage == "ZIN_ind"][cohort < 55]
downy_dates[primary_infection_stage == "SUZ_death_ind"][cohort < 55]

downy_dates[primary_infection_stage == "INC_h_upper"][cohort < 55]
```



## 2024 Season

### Experimental design  



### Spray applications

 - 2024-09-04  Miravis  
 - 2024-09-18  Captan  
 - 2024-10-02  EcoCarb Sulphur mix  

